A weakly structured stem for human origins in Africa

### 2022-08-12 paper.tex
point to validations in supplement

Regardless of the model choice for early epochs, maximum-likelihood inference of human demographic
history for the last 150ka was remarkably robust.
In a reticulated model, we use ``divergence'' between populations to mean
the time of their most recent shared ancestry.
The earliest divergence among
contemporary human populations differentiates the southern African Nama from
other African groups between 110--135ka, with low to moderate levels of
subsequent gene flow (Table \ref{tab:migration-rates}). In none of the
high-likelihood models which we explored did the divergence between Nama and
other populations exceed $\sim$140ka. 
We conclude that geographic patterns of contemporary \emph{Homo sapiens} population structure
likely arise during MIS 5.
Although we find evidence for earlier population structure in Africa (see
below), contemporary populations cannot be easily mapped onto the more ancient
`stem' groups as only a small proportion of drift between contemporary populations can
be attributed to drift between stems (Section \ref{sec:f4},  Figures~\ref{fig:predictions} and
\ref{fig:supp-f4s-single-origin}--\ref{fig:supp-f4s-merger-with-stem-migration}).

>>>>

Regardless of the model choice for early epochs, maximum-likelihood inference of human demographic
history for the last 150ka was remarkably robust.
In a reticulated model, we use ``divergence'' between populations to mean
the time of their most recent shared ancestry.
The earliest divergence among
contemporary human populations differentiates the southern African Nama from
other African groups between 110--135ka, with low to moderate levels of
subsequent gene flow (Table \ref{tab:migration-rates}). In none of the
high-likelihood models which we explored did the divergence between Nama and
other populations exceed $\sim$140ka
(Supp. Information~\ref{sec:validation-data-models} describes validation using
alternative datasets and model specifications). 
We conclude that geographic patterns of contemporary \emph{Homo sapiens}
population structure likely arise during MIS 5.
Although we find evidence for earlier population structure in Africa (next section),
contemporary populations cannot be easily mapped onto the more ancient
`stem' groups as only a small proportion of drift between contemporary populations can
be attributed to drift between stems (Section \ref{sec:f4},  Figures~\ref{fig:predictions} and
\ref{fig:supp-f4s-single-origin}--\ref{fig:supp-f4s-merger-with-stem-migration}).
